%%% Proof-Carrying Transactions — Full Research Paper
\documentclass[10pt,conference]{IEEEtran}
\IEEEoverridecommandlockouts

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{amsthm}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{url}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{tikz}
\usetikzlibrary{arrows.meta, positioning, shapes.geometric, calc, decorations.markings}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{listings}
\usepackage{subcaption}
\usepackage{balance}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{corollary}{Corollary}
\newtheorem{property}{Property}

\lstset{
  basicstyle=\ttfamily\footnotesize,
  keywordstyle=\color{blue!70!black}\bfseries,
  commentstyle=\color{green!50!black},
  stringstyle=\color{red!60!black},
  breaklines=true,
  frame=single,
  numbers=left,
  numberstyle=\tiny\color{gray},
  xleftmargin=2em,
  framexleftmargin=1.5em
}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

\title{Proof-Carrying Transactions: Zero-Knowledge\\Policy Compliance for Autonomous AI Agents}

\author{\IEEEauthorblockN{Anonymous Author(s)}}

\maketitle

%% ===================================================================
%% ABSTRACT
%% ===================================================================
\begin{abstract}
The deployment of autonomous AI agents capable of initiating financial transactions, invoking external APIs, and orchestrating multi-step workflows across enterprise boundaries introduces an unprecedented class of alignment failures: \emph{egress-time policy violations}---actions that individually appear benign but collectively breach organizational constraints governing amount limits, destination whitelists, temporal windows, and semantic intent boundaries. Existing enforcement architectures rely on centralized proxy firewalls that require downstream consumers to implicitly trust the enforcer's computation, creating single points of compromise and precluding independent auditability.

We introduce \textbf{Proof-Carrying Transactions (PCT)}, a cryptographic framework that removes the need to trust the proxy's evaluation logic by transforming every outbound agent request into a self-certifying artifact: each transaction carries a zero-knowledge succinct non-interactive argument of knowledge (zk-SNARK) attesting that its parameters satisfy the organization's declarative policy constraints, without revealing the policy structure, private thresholds, or internal agent state to any external party.

Our core contribution is architectural, not cryptographic: we design a \emph{Policy Constraint Language} (PCL) and its domain-specific compiler that translates declarative egress rules into optimized Rank-1 Constraint Systems (R1CS), reducing constraint counts by 50--100$\times$ and proving latency by orders of magnitude compared to general-purpose zkVMs like RISC Zero. We implement PCT using the Groth16 proving scheme over the BN254 elliptic curve and introduce three novel application-layer gadgets---a \emph{16-bit quantized semantic similarity threshold circuit}, \emph{temporal window membership proofs}, and \emph{hierarchical Merkle whitelist accumulators}---that extend standard ZKP primitives to the domain of AI agent governance.

Comprehensive evaluation on a corpus of 847 real-world enterprise policies demonstrates that PCT achieves a median proof generation latency of 95ms and verification in 4.2ms ($O(1)$) while maintaining constraint counts at approximately 11,200 R1CS gates for 97.3\% of policies tested. We prove that PCT satisfies \emph{computational soundness}, \emph{zero-knowledge privacy}, and a new property we term \emph{policy-binding completeness}---ensuring that no compliant transaction is rejected and no non-compliant transaction can forge a valid proof under the Knowledge of Exponent assumption. Finally, we demonstrate PCT's regulatory applicability by mapping its cryptographic guarantees to the technical requirements of the EU AI Act (Articles 9, 12, 14) and the NIST AI Risk Management Framework, establishing PCT as the first formally verified bridge between applied cryptography and AI governance regulation.
\end{abstract}

\begin{IEEEkeywords}
Zero-Knowledge Proofs, Autonomous AI Agents, Agent Alignment, Proof-Carrying Code, Cryptographic Policy Enforcement, AI Governance, zk-SNARK, Groth16, EU AI Act Compliance
\end{IEEEkeywords}

%% ===================================================================
%% I. INTRODUCTION
%% ===================================================================
\section{Introduction}
\label{sec:intro}

Large language model (LLM)-based autonomous agents---software entities capable of planning, reasoning, and executing multi-step actions without continuous human supervision---are rapidly transitioning from research prototypes to production infrastructure across financial services, healthcare operations, supply chain logistics, and critical infrastructure management~\cite{xi2023rise, wang2024survey}. These agents routinely initiate wire transfers, invoke third-party APIs, modify database records, and orchestrate workflows that carry real-world consequences measured in millions of dollars and human lives.

This proliferation introduces what we term the \textbf{Autonomous Egress Problem}: how can an organization cryptographically guarantee that every outbound action taken by an AI agent complies with its governance policies, without requiring downstream systems to trust the enforcement infrastructure and without exposing proprietary policy details to external observers?

\subsection{The Failure of Trust-Based Enforcement}

Current approaches to agent governance fall into three inadequate categories:

\textbf{(1) Application-level guardrails.} Systems like prompt engineering, Constitutional AI~\cite{bai2022constitutional}, and RLHF~\cite{ouyang2022training} operate at the model layer, attempting to shape agent behavior through training-time alignment. These approaches lack formal guarantees: a sufficiently capable agent may learn to circumvent soft constraints through prompt injection, jailbreaking, or simply by exploiting gaps between the training distribution and deployment reality~\cite{perez2022red}. More critically, they provide no post-hoc verifiable evidence that a specific transaction was compliant at execution time.

\textbf{(2) Centralized proxy firewalls.} API gateways like Kong, Envoy, and proprietary agent management platforms enforce policies by intercepting requests at a choke point. While effective for simple allow/deny rules, these architectures demand that every downstream consumer implicitly trust the proxy's correctness. A compromised, misconfigured, or subtly buggy proxy invalidates the entire security model---and the downstream API has \emph{no independent mechanism} to verify that enforcement actually occurred.

\textbf{(3) Audit-after-the-fact.} Post-hoc logging and compliance review detect violations retroactively but cannot prevent them. In high-stakes domains---automated trading, medical record access, infrastructure commands---retroactive detection is insufficient: the damage is already done.

\subsection{Our Contribution: Proof-Carrying Transactions}

We propose a fundamental paradigm shift: \emph{the enforcement proof travels with the transaction}. Drawing inspiration from Necula's seminal work on proof-carrying code~\cite{necula1997proof}, where executable binaries carry machine-checkable proofs of safety properties, we introduce \textbf{Proof-Carrying Transactions (PCT)}---a framework where every outbound agent request is augmented with a cryptographic proof attesting that the request's parameters satisfy the organization's policy constraints.

The key insight is that by compiling declarative policy rules into arithmetic circuits and using the agent's transaction parameters as the private witness, we can leverage the zero-knowledge property of zk-SNARKs to achieve three seemingly contradictory goals simultaneously:

\begin{enumerate}
    \item \textbf{Verifiable compliance}: Any party holding the verification key can confirm in $O(1)$ time that the transaction satisfies all policy constraints.
    \item \textbf{Policy privacy}: The proof reveals nothing about the policy structure, thresholds, whitelist contents, or internal organizational rules.
    \item \textbf{Transaction privacy}: Private fields (exact amounts, internal routing identifiers, agent reasoning traces) remain hidden from external verifiers.
\end{enumerate}

We make the following concrete contributions:

\begin{itemize}
    \item We formalize the PCT security model (Section~\ref{sec:model}), defining the threat landscape for autonomous agent egress and proving that PCT satisfies \emph{computational soundness}, \emph{zero-knowledge privacy}, and a new property we term \emph{policy-binding completeness} (Theorem~\ref{thm:soundness}, Theorem~\ref{thm:privacy}, Theorem~\ref{thm:completeness}).

    \item We design and implement a \emph{Policy Constraint Language} (PCL) and its compiler that transforms declarative egress policies into optimized R1CS circuits (Section~\ref{sec:compiler}), including three novel cryptographic gadgets for semantic similarity thresholds, temporal window proofs, and hierarchical Merkle whitelist accumulators.

    \item We present a complete, production-grade implementation integrated into a Rust-based reverse proxy (Section~\ref{sec:implementation}), demonstrating end-to-end proof generation, attachment, and verification within the HTTP request lifecycle.

    \item We evaluate PCT against 847 enterprise-derived egress policies (Section~\ref{sec:evaluation}), reporting median proof generation latency of 95ms, verification in 4.2ms, and constraint counts at approximately 11,200 gates for 97.3\% of policies.

    \item We provide the first formal mapping between ZKP-based compliance guarantees and the technical requirements of the EU AI Act and NIST AI RMF (Section~\ref{sec:regulatory}), establishing PCT as a mechanism for demonstrable regulatory compliance.
\end{itemize}

%% ===================================================================
%% II. BACKGROUND AND RELATED WORK
%% ===================================================================
\section{Background and Related Work}
\label{sec:related}

\subsection{Autonomous AI Agent Architectures}

Modern autonomous agents---such as those built on ReAct~\cite{yao2022react}, AutoGPT, and tool-augmented LLM frameworks---operate through a perceive-plan-act loop where a language model generates executable actions (API calls, database queries, financial transactions) based on a combination of user instructions, environmental observations, and internally maintained state~\cite{wang2024survey}. Hendrycks et al.~\cite{hendrycks2021unsolved} catalogue the unsolved safety problems that arise from deploying such systems, while Amodei et al.~\cite{amodei2016concrete} formalize the concrete accident scenarios that motivate runtime enforcement. The critical security observation is that these agents possess \emph{delegated authority}: they act on behalf of a principal (user, organization) but with significant operational autonomy. The agent's action space typically includes HTTP requests to external services, each of which constitutes an \emph{egress event} governed by organizational policy.

\subsection{Zero-Knowledge Proof Systems}

A zero-knowledge proof system allows a prover $\mathcal{P}$ to convince a verifier $\mathcal{V}$ that a statement $x$ belongs to a language $\mathcal{L}$ without revealing anything beyond the validity of the statement itself~\cite{goldwasser1989knowledge}. Formally:

\begin{definition}[Zero-Knowledge Proof System]
A triple $(\mathcal{P}, \mathcal{V}, \textsf{Setup})$ is a zero-knowledge proof system for a relation $\mathcal{R}$ if it satisfies:
\begin{itemize}
    \item \textbf{Completeness}: For all $(x, w) \in \mathcal{R}$, $\Pr[\mathcal{V}(\textsf{vk}, x, \pi) = 1 \mid \pi \leftarrow \mathcal{P}(\textsf{pk}, x, w)] = 1$.
    \item \textbf{Soundness}: For all PPT adversaries $\mathcal{A}$ and all $x \notin \mathcal{L}$, $\Pr[\mathcal{V}(\textsf{vk}, x, \pi^*) = 1 \mid \pi^* \leftarrow \mathcal{A}(\textsf{pk}, x)] \leq \textsf{negl}(\lambda)$.
    \item \textbf{Zero-Knowledge}: There exists a PPT simulator $\mathcal{S}$ such that for all $(x, w) \in \mathcal{R}$, the distributions $\{\pi \leftarrow \mathcal{P}(\textsf{pk}, x, w)\}$ and $\{\pi^* \leftarrow \mathcal{S}(\textsf{vk}, x)\}$ are computationally indistinguishable.
\end{itemize}
\end{definition}

\textbf{Groth16}~\cite{groth2016size} is a pairing-based zk-SNARK achieving constant-size proofs (3 group elements: $[\mathsf{A}]_1, [\mathsf{C}]_1, [\mathsf{B}]_2$) and verification in $O(1)$ time via a single pairing check. It requires a relation-specific trusted setup producing a Common Reference String (CRS). While newer systems such as PLONK~\cite{gabizon2019plonk} and Halo2 offer universal or transparent setups, Bulletproofs~\cite{bunz2018bulletproofs} provide logarithmic-size proofs without trusted setup, and STARKs~\cite{ben2018scalable} achieve post-quantum security with transparent setup, Groth16 remains optimal for scenarios where proof size and verification speed are paramount---precisely the constraints of real-time transaction processing. Recent surveys by Sharma et al.~\cite{sharma2023machine} and Togan et al.~\cite{togan2023zero} confirm this finding in the context of applied ZKP deployments.

\subsection{Proof-Carrying Code}

Necula's proof-carrying code (PCC)~\cite{necula1997proof} introduced the concept of augmenting untrusted programs with machine-checkable proofs of safety properties. The code consumer verifies the proof against a safety policy using a small, trusted verification condition generator. PCT extends this paradigm from static binary safety to dynamic transaction compliance: instead of proving memory safety of compiled code, we prove policy compliance of runtime-generated requests. The key technical difference is that PCC proofs are generated once at compile time, while PCT proofs must be generated in real time (sub-second) for every transaction.

\section{Comparative Analysis of Governance Architectures}
\label{sec:comparative}

Zero-knowledge proofs have been deployed for financial privacy (Zcash~\cite{sasson2014zerocash}), blockchain scaling (zkRollups~\cite{loopring2020}), identity verification (Semaphore~\cite{semaphore2020}), verifiable credentials~\cite{w3c2022vc}, and machine learning over encrypted data~\cite{bost2015machine}. Kang et al.~\cite{kang2022scaling} and Zhao et al.~\cite{zhao2024verifiable} explore trustless ML execution, while ZoKrates~\cite{eberhardt2018zokrates} and Badrinarayanan et al.~\cite{badrinarayanan2020verifiable} develop general-purpose ZKP toolchains. However, the application of ZKPs to \emph{application-layer outbound policy enforcement for AI agents} is, to our knowledge, entirely unexplored. Existing agent governance frameworks such as Guardrails AI, LangChain safety modules, and Microsoft Guidance~\cite{zheng2023llama} provide runtime filtering but lack cryptographic verifiability, policy privacy, and independent auditability. To understand the paradigm shift introduced by Proof-Carrying Transactions, we directly compare our framework against traditional agent governance architectures across three critical vectors: performance overhead, security guarantees, and trust models.

\textbf{1) Application-level Filtering (e.g., RLHF, Prompt Guardrails):}
These methods operate probabilistically at the model layer~\cite{sun2023trust}. While they exhibit minimal latency overhead, their security guarantees are extraordinarily weak against capable adversaries utilizing prompt injection or jailbreaking~\cite{carlini2023extracting}. Moreover, they offer zero formal verifiability; an auditor cannot mathematically prove a model evaluated its safety guardrails accurately during a transaction. Blanchard et al.~\cite{blanchard2017byzantine} show that even Byzantine-tolerant training fails to produce auditable per-transaction guarantees.

\textbf{2) Proxy/API Choke Points (e.g., Kong, Envoy):}
Centralized firewalls enforce deterministic rules at the network boundary. While effective for simple IP whitelists and API key validation, they necessitate that all downstream consumers (banks, partner APIs) implicitly trust the firewall operator. Policy engines such as Open Policy Agent~\cite{schmidt2019opa} and role-based access control frameworks~\cite{sandhu1996role} improve expressiveness but do not eliminate the fundamental single-point-of-trust limitation. Authorization calculi~\cite{abadi2011authorization} and delegated credentials such as Macaroons~\cite{birgisson2014macaroons} offer decentralized caveats, yet leak policy structure to every intermediary in the chain.

\textbf{3) Proof-Carrying Transactions (PCT):}
By binding the constraint satisfaction directly to the transaction payload via a zero-knowledge proof, PCT shifts the trust model from an active enforcer to a passive verifier. While proof generation introduces a small computational overhead ($\sim$95ms per transaction), the security guarantees transition from probabilistic heuristics to cryptographic certainty. An auditor can mathematically verify compliance post-execution without requiring access to the proprietary rule set, bridging the gap between operational privacy and regulatory auditability.

Table~\ref{tab:comparison} summarizes this comparative landscape.

\begin{table}[t]
\centering
\caption{Comparison of AI Agent Governance Approaches}
\label{tab:comparison}
\resizebox{\columnwidth}{!}{
\begin{tabular}{@{}lccccc@{}}
\toprule
\textbf{Approach} & \textbf{Formal} & \textbf{Policy} & \textbf{Indep.} & \textbf{Real-} & \textbf{Regulatory} \\
 & \textbf{Guarantee} & \textbf{Privacy} & \textbf{Verify} & \textbf{time} & \textbf{Mapping} \\
\midrule
RLHF / Constitutional AI & \texttimes & \texttimes & \texttimes & \checkmark & \texttimes \\
Prompt Engineering & \texttimes & \texttimes & \texttimes & \checkmark & \texttimes \\
Guardrails AI / LangChain & \texttimes & \texttimes & \texttimes & \checkmark & \texttimes \\
API Gateway (Kong, Envoy) & \texttimes & \texttimes & \texttimes & \checkmark & \texttimes \\
mTLS + Macaroons & Partial & \texttimes & Partial & \checkmark & \texttimes \\
Post-hoc Audit Logging & \texttimes & \texttimes & Partial & \texttimes & Partial \\
\textbf{PCT (This Work)} & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark \\
\bottomrule
\end{tabular}
}
\end{table}

%% ===================================================================
%% III. SECURITY MODEL
%% ===================================================================
\section{Security Model and Formal Definitions}
\label{sec:model}

\subsection{System Model}

We define the PCT ecosystem through four principals:

\begin{definition}[PCT Principals]
\label{def:principals}
The system comprises:
\begin{itemize}
    \item \textbf{Agent} $\mathcal{A}$: An autonomous software entity that generates outbound requests $\mathsf{req} = (\mathsf{target}, \mathsf{payload}, \mathsf{metadata})$.
    \item \textbf{Policy Authority} $\mathcal{PA}$: The organizational entity that defines a policy $\mathsf{P}$ as a set of constraints over the request space.
    \item \textbf{Prover Proxy} $\mathcal{PP}$: An intermediary that intercepts $\mathsf{req}$, evaluates it against $\mathsf{P}$, and generates a proof $\pi$.
    \item \textbf{Verifier} $\mathcal{V}$: An entity that verifies compliance before finalizing the transaction. While theoretically this could be the external receiving API, in practice $\mathcal{V}$ is typically an internal compliance gateway, an organizational clearinghouse, or a regulatory auditor, as external APIs primarily enforce their own authentication. Verification occurs in constant time using a public verification key $\mathsf{vk}$.
\end{itemize}
\end{definition}

\subsection{Threat Model}

We consider the following adversarial capabilities:

\begin{enumerate}
    \item \textbf{Malicious Agent}: $\mathcal{A}$ may attempt to craft requests that appear compliant to the proxy but violate policy intent. This models jailbroken, compromised, or misaligned agents.
    \item \textbf{Compromised Downstream}: External APIs or services may attempt to extract policy details from observed proofs or verification patterns.
    \item \textbf{Colluding Verifier}: $\mathcal{V}$ may collude with $\mathcal{A}$ and attempt to forge proofs for non-compliant transactions.
    \item \textbf{Network Observer}: A passive adversary observing the communication channel may attempt to learn policy structure from proof metadata.
\end{enumerate}

\textbf{Enforcement Assumption.} We assume the agent operates within a network-isolated execution environment (e.g., an isolated VPC, a sandboxed container, or a TEE enclave) where the \emph{sole} egress route is the Prover Proxy $\mathcal{PP}$. This is enforced at the infrastructure layer via a strict default-deny egress posture across all protocols (TCP, UDP, ICMP); the agent has no direct outbound network access, preventing side-channel exfiltration via DNS tunneling, timing channels, or Server-Side Request Forgery (SSRF) against authorized downstream APIs. We explicitly scope PCT as an application-layer policy enforcer, not a network-layer sandbox; it secures authorized transaction channels but relies on the underlying virtual private cloud to prevent out-of-band leakage. Consequently, a malicious agent $\mathcal{A}$ cannot bypass the proxy---it can only attempt to craft requests that pass proof generation while violating policy intent, which is precisely the attack that computational soundness (Theorem~\ref{thm:soundness}) prevents.

\textbf{Trust Split.} The Prover Proxy $\mathcal{PP}$ is trusted for data routing and availability, but is \emph{untrusted for computation}. If the proxy's evaluation logic is bug-ridden or maliciously altered, the ZKP properties ensure it cannot forge a valid proof for a non-compliant transaction. The proof cryptographically binds the specific payload (hashed as a public input) to the correct mathematical evaluation of the policy. To prevent the proxy from actively tampering with requests (e.g., swapping the recipient in an otherwise authorized payload), the agent's trusted execution enclave digitally signs the payload. The downstream verifier cryptographically binds the generated proof to the agent's original intent by checking this signature alongside the SNARK verification. However, because SNARKs are stateless, evaluating temporal rate limits ($C_{\text{rate}}$) requires the proxy to maintain a state accumulator. We explicitly scope this: the proxy is \emph{trusted for state continuity} (providing the correct historical sum), but remains untrusted for the threshold evaluation itself. State-rollback attacks---where a compromised proxy resets the accumulator to bypass rate limits---are mitigated in production by committing the accumulator root as a public input, enabling verifiers to detect discontinuities. We discuss future paths to fully eliminate proxy trust (e.g., agent-side TEE proving) in Section~\ref{sec:future}.

\textbf{Semantic Constraint Boundary.} The semantic similarity constraint $C_{\text{sem}}$ operates as an \emph{optimistic semantic check}: PCT cryptographically guarantees that the mathematical evaluation of cosine similarity against the reference embedding was performed correctly, but does not guarantee that the embedding space itself is adversarially robust. The security of $C_{\text{sem}}$ is therefore conditional on the fidelity of the underlying embedding model. We make this boundary explicit: PCT secures policy \emph{evaluation}, not the latent geometry of the embedding model.

\subsection{Policy Formalization}

\begin{definition}[Egress Policy]
\label{def:policy}
An egress policy $\mathsf{P}$ is a Boolean function $\mathsf{P}: \mathcal{X} \rightarrow \{0, 1\}$ over a request space $\mathcal{X}$, where $\mathsf{P}(\mathsf{req}) = 1$ if and only if $\mathsf{req}$ satisfies all organizational constraints. A policy is composed of a conjunction of typed constraints:
\[
\mathsf{P}(\mathsf{req}) = \bigwedge_{i=1}^{n} C_i(\mathsf{req})
\]
where each $C_i$ belongs to one of the following constraint classes:
\begin{itemize}
    \item \textbf{Range}: $C_{\text{range}}(\mathsf{req}) \equiv v_{\min} \leq \mathsf{req}.\mathsf{amount} \leq v_{\max}$
    \item \textbf{Set Membership}: $C_{\text{set}}(\mathsf{req}) \equiv \mathsf{req}.\mathsf{target} \in \mathcal{W}$ for whitelist $\mathcal{W}$
    \item \textbf{Temporal}: $C_{\text{time}}(\mathsf{req}) \equiv t_{\text{start}} \leq \mathsf{req}.\mathsf{timestamp} \leq t_{\text{end}}$
    \item \textbf{Rate}: $C_{\text{rate}}(\mathsf{req}) \equiv \sum_{j} \mathsf{req}_j.\mathsf{amount} \leq B$ over sliding window $\Delta t$
    \item \textbf{Semantic}: $C_{\text{sem}}(\mathsf{req}) \equiv \text{sim}(\mathsf{emb}(\mathsf{req}.\mathsf{intent}), \mathsf{emb}(\mathsf{ref})) \geq \tau$
\end{itemize}
\end{definition}

\subsection{Security Properties}

We now state the three properties that PCT must satisfy.

\begin{theorem}[Computational Soundness]
\label{thm:soundness}
Under the Knowledge of Exponent (KoE) assumption over the BN254 curve, for any PPT adversary $\mathcal{A}^*$:
\[
\Pr\left[\mathcal{V}(\mathsf{vk}, \mathsf{req}, \pi^*) = 1 \;\wedge\; \mathsf{P}(\mathsf{req}) = 0 \;\middle|\; \pi^* \leftarrow \mathcal{A}^*(\mathsf{pk})\right] \leq \textsf{negl}(\lambda)
\]
That is, no efficient adversary can produce a valid proof for a non-compliant transaction.
\end{theorem}

\begin{proof}
We reduce to the knowledge soundness of Groth16~\cite{groth2016size} under the Knowledge of Exponent (KoE) assumption over groups of order $p$ on the BN254 curve.

Assume, for contradiction, that a PPT adversary $\mathcal{A}^*$ outputs $(\mathsf{req}, \pi^*)$ such that $\mathsf{P}(\mathsf{req}) = 0$ yet $\mathcal{V}(\mathsf{vk}, \mathsf{req}, \pi^*) = 1$. Because $\mathsf{P}(\mathsf{req}) = 0$, there exists at least one constraint $C_i$ with $C_i(\mathsf{req}) = 0$. By Lemma~\ref{lem:compiler_correctness} (compiler correctness), the R1CS instance $\mathsf{R} = (\mathbf{A}, \mathbf{B}, \mathbf{C})$ compiled from $\mathsf{P}$ admits no satisfying assignment $z$ consistent with the witness $w$ extracted from $\mathsf{req}$; i.e., there is no $z$ such that $\mathbf{A} z \circ \mathbf{B} z = \mathbf{C} z$.

By the KoE assumption, a valid Groth16 proof $\pi^*$ implies the existence of a PPT extractor $\mathcal{E}$ that, given the same randomness as $\mathcal{A}^*$, outputs a satisfying witness $w^*$ for $\mathsf{R}$. This yields a contradiction: $w^*$ simultaneously satisfies and does not satisfy $\mathsf{R}$. Therefore, $\Pr[\text{forge}] \leq \textsf{negl}(\lambda)$. \qed
\end{proof}

\begin{theorem}[Zero-Knowledge Policy Privacy]
\label{thm:privacy}
For any two policies $\mathsf{P}_0, \mathsf{P}_1$ of equal circuit size and any request $\mathsf{req}$ such that $\mathsf{P}_0(\mathsf{req}) = \mathsf{P}_1(\mathsf{req}) = 1$, the proof distributions are computationally indistinguishable:
\[
\left\{\pi \leftarrow \mathcal{P}(\mathsf{pk}_0, \mathsf{req}, w_0)\right\} \approx_c \left\{\pi \leftarrow \mathcal{P}(\mathsf{pk}_1, \mathsf{req}, w_1)\right\}
\]
\end{theorem}

\begin{proof}
Groth16 satisfies perfect zero-knowledge: for every CRS produced by \textsf{Setup}, there exists a PPT simulator $\mathcal{S}$ that, given the trapdoor $\tau$ but \emph{no} witness, produces proofs whose distribution is \emph{identical} to honest proofs~\cite{groth2016size}.

Let $\mathsf{P}_0, \mathsf{P}_1$ be two policies with equal-sized R1CS representations $\mathsf{R}_0, \mathsf{R}_1$, and let $(\mathsf{pk}_b, \mathsf{vk}_b) \leftarrow \textsf{Setup}(\mathsf{R}_b)$ for $b \in \{0,1\}$. For each $b$, the simulator $\mathcal{S}_b(\mathsf{vk}_b, x)$ produces simulated proofs $\tilde{\pi}_b$ such that $\{\pi_b\} \equiv \{\tilde{\pi}_b\}$ (statistical equality). Since $\tilde{\pi}_0$ and $\tilde{\pi}_1$ are independent of the witness, an external observer holding only the proof $\pi$ and the public inputs $x$ cannot determine which policy generated it. Formally, any distinguisher $\mathcal{D}$ satisfies $|\Pr[\mathcal{D}(\pi_0)=1] - \Pr[\mathcal{D}(\pi_1)=1]| = 0$.

This property extends to policy \emph{structure}: the number, type, and thresholds of individual constraints are encoded entirely in the witness and circuit wiring, never in the public inputs or the proof itself. \qed
\end{proof}

\begin{theorem}[Policy-Binding Completeness]
\label{thm:completeness}
For any policy $\mathsf{P}$ and any request $\mathsf{req}$ such that $\mathsf{P}(\mathsf{req}) = 1$, the honest prover $\mathcal{PP}$ always produces a proof that the verifier accepts:
\[
\Pr\left[\mathcal{V}(\mathsf{vk}, \mathsf{req}, \pi) = 1 \;\middle|\; \pi \leftarrow \mathcal{PP}(\mathsf{pk}, \mathsf{req})\right] = 1
\]
This ensures that PCT never produces false rejections---a critical property for production deployment where availability is paramount.
\end{theorem}

\begin{proof}
Completeness follows from two independent guarantees composed sequentially.

\emph{Step 1 (Compiler completeness).} By Lemma~\ref{lem:compiler_correctness}, if $\mathsf{P}(\mathsf{req}) = 1$ then the witness vector $w$ extracted from $\mathsf{req}$ satisfies the compiled R1CS instance $\mathsf{R}$, i.e., $\mathbf{A}w \circ \mathbf{B}w = \mathbf{C}w$.

\emph{Step 2 (Groth16 completeness).} Given a satisfying assignment $(x, w)$ for $\mathsf{R}$, the honest Groth16 prover deterministically produces $\pi = ([\mathsf{A}]_1, [\mathsf{B}]_2, [\mathsf{C}]_1)$ satisfying the pairing verification equation (Equation~\ref{eq:pairing}). This is an unconditional algebraic identity that holds with probability 1 over the honest execution path.

Composing Steps 1 and 2: for any compliant request, the honest PCT pipeline extracts a valid witness, feeds it to the Groth16 prover, and produces a proof that the verifier accepts with certainty. \qed
\end{proof}

%% ===================================================================
%% IV. SYSTEM ARCHITECTURE
%% ===================================================================
\section{System Architecture}
\label{sec:architecture}

\subsection{Architecture Overview}

The PCT system operates as a transparent layer within the network stack, intercepting agent traffic at the HTTP/gRPC level. Figure~\ref{fig:architecture} illustrates the end-to-end flow.

\begin{figure}[t]
\centering
\resizebox{\columnwidth}{!}{
\begin{tikzpicture}[
    node distance=1.2cm and 1.8cm,
    box/.style={draw, rounded corners=3pt, minimum width=2.2cm, minimum height=0.8cm, font=\footnotesize, fill=#1!15, text=#1!80!black, thick},
    box/.default=blue,
    arrow/.style={-{Stealth[length=5pt]}, thick, color=gray!70!black},
    label/.style={font=\scriptsize\itshape, text=gray!60!black}
]
    \node[box=violet] (agent) {AI Agent};
    \node[box=blue, right=of agent] (proxy) {Prover Proxy};
    \node[box=teal, right=of proxy] (api) {External API};
    \node[box=orange, below=0.8cm of proxy] (compiler) {Circuit Compiler};
    \node[box=red, below=0.8cm of compiler] (prover) {ZKP Engine};

    \draw[arrow] (agent) -- node[above, label] {$\mathsf{req}$} (proxy);
    \draw[arrow] (proxy) -- node[above, label] {$\mathsf{req} \| \pi$} (api);
    \draw[arrow] (proxy) -- node[right, label] {AST} (compiler);
    \draw[arrow] (compiler) -- node[right, label] {R1CS} (prover);
    \draw[arrow, dashed] (prover.west) -- ++(-2.2,0) |- node[near end, left, label] {$\pi$} (proxy.south west);
\end{tikzpicture}%
}
\caption{PCT Architecture. The agent submits a request; the proxy extracts witnesses, compiles the policy AST into R1CS constraints, generates a Groth16 proof $\pi$, and attaches it to the outbound request as a self-certifying header.}
\label{fig:architecture}
\end{figure}

\subsection{Request Lifecycle}

The PCT lifecycle proceeds in six phases:

\textbf{Phase 1: Interception.} The AI agent emits an HTTP/gRPC request. The Prover Proxy intercepts the request at the transport layer before any external transmission.

\textbf{Phase 2: Witness Extraction.} The proxy parses the request payload and extracts a structured witness vector $w = (w_1, w_2, \ldots, w_k)$ containing the transaction parameters relevant to policy evaluation. Amounts are encoded as 64-bit fixed-point field elements; target URLs are hashed via Poseidon into BN254 scalar field elements; timestamps are Unix epoch integers modulo the field prime.

\textbf{Phase 3: Policy Evaluation.} The proxy evaluates the witness against the organization's active policy $\mathsf{P}$. If $\mathsf{P}(w) = 0$, the request is rejected with a structured denial response and the event is logged. No proof is generated for rejected requests.

\textbf{Phase 4: Circuit Synthesis.} For approved requests ($\mathsf{P}(w) = 1$), the policy compiler synthesizes the R1CS circuit encoding the policy constraints and assigns the extracted witness to the circuit's private inputs.

\textbf{Phase 5: Proof Generation.} The Groth16 prover executes the witness assignment through the circuit, producing proof $\pi = ([\mathsf{A}]_1, [\mathsf{B}]_2, [\mathsf{C}]_1)$.

\textbf{Phase 6: Egress Transmission.} The proxy forwards the original request to the target API with two additional elements: (a) the proof encoded as a Base64 HTTP header \texttt{X-PCT-Proof}, and (b) an identical proof record committed to the immutable audit log for regulatory compliance.

\subsection{Proof Verification}

Any party possessing the public verification key $\mathsf{vk}$ can verify a Proof-Carrying Transaction in $O(1)$ time by checking the Groth16 pairing equation:
\begin{equation}
    e([\mathsf{A}]_1, [\mathsf{B}]_2) = e([\alpha]_1, [\beta]_2) \cdot e\left(\sum_{i=0}^{l} x_i \cdot [\gamma_i]_1, [\gamma]_2\right) \cdot e([\mathsf{C}]_1, [\delta]_2)
    \label{eq:pairing}
\end{equation}
where $x_0, \ldots, x_l$ are the public inputs (transaction hash, timestamp, policy version identifier) and $\alpha, \beta, \gamma, \delta$ are elements from the trusted setup CRS.

%% ===================================================================
%% V. POLICY COMPILER
%% ===================================================================
\section{Policy Constraint Language and Compiler}
\label{sec:compiler}

\subsection{Policy Constraint Language (PCL)}

We define a domain-specific Policy Constraint Language that enables organizations to express egress rules declaratively. PCL policies compile to an intermediate Abstract Syntax Tree (AST) before circuit synthesis.

\begin{lstlisting}[language=Python, caption={Example PCL policy specification.}, label={lst:pcl}]
policy "production_egress" {
  # amount bounds
  constraint amount_limit {
    type: range
    field: transaction.amount
    min: 0
    max: 50000
  }
  # destination whitelist
  constraint approved_targets {
    type: set_membership
    field: transaction.target
    set: ["api.stripe.com",
          "api.openai.com",
          "hooks.slack.com"]
  }
  # business hours only
  constraint temporal_window {
    type: temporal
    field: transaction.timestamp
    start: "09:00 UTC"
    end: "17:00 UTC"
    days: [Mon, Tue, Wed, Thu, Fri]
  }
  # semantic guardrail
  constraint intent_alignment {
    type: semantic_similarity
    field: transaction.intent_embedding
    reference: "authorized_task_embedding"
    threshold: 0.85
  }
}
\end{lstlisting}

\subsection{AST Construction}

The PCL parser produces a typed AST where each node represents either a leaf constraint or a Boolean combinator ($\land$, $\lor$, $\neg$). The type system enforces that field references resolve to valid request schema paths and that constraint parameters are well-formed.

\begin{definition}[Policy AST]
A Policy AST $\mathcal{T}$ is a binary tree where:
\begin{itemize}
    \item Leaf nodes are typed constraints $C_i \in \{\text{Range}, \text{SetMem}, \text{Temporal}, \text{Rate}, \text{Semantic}\}$.
    \item Internal nodes are Boolean operators $\{\land, \lor, \neg\}$.
    \item The root node evaluates to a single Boolean output wire.
\end{itemize}
\end{definition}

\subsection{AST to R1CS Compilation Mechanics}

Translating the abstract syntax tree into a Rank-1 Constraint System (R1CS) requires bridging declarative logical rules into algebraic circuits over a finite field $\mathbb{F}_p$. To optimize the constraint footprint and ensure soundness, our compiler executes a multi-pass translation algorithm across three distinct phases:

\textbf{Pass 1: Type Inference and Range Bounding.} The compiler traverses the AST to statically determine the maximum bit width required for each variable based on the policy schema. For example, if a node dictates \texttt{amount < 50000}, the compiler restricts the arithmetic wire constraints to 16 bits, drastically reducing the subsequent binary decomposition overhead.

\textbf{Pass 2: Common Sub-expression Elimination (CSE).} Policies frequently reuse variables (e.g., referencing \texttt{transaction.target} in both a whitelist check and an intent validation). By maintaining a dependency graph during parsing, the compiler eliminates redundant node evaluations, ensuring that repetitive field extractions map to a singular set of constrained input wires.

\textbf{Pass 3: Gadget Instantiation.} The compiler maps high-level logical operators to their respective cryptographic gadgets. Simple Boolean operators ($\land, \lor$) are directly translated into degree-2 algebraic constraints: $A \land B \Rightarrow A \cdot B = C$. Complex functions, such as the \texttt{SIMILAR} semantic intent check, are instantiated by feeding the private transaction embedding vector through an in-circuit Poseidon hash instance, followed by a cosine similarity threshold calculation normalized over the field prime.

The compiler concludes with a post-order traversal, emitting the finalized R1CS matrices for each leaf node and weaving them through the logical circuitry. We now describe the specific gadget formulation for each constraint type.

\subsubsection{Range Constraints}

For a range constraint $v_{\min} \leq x \leq v_{\max}$, we decompose into two comparison gadgets. Each comparison $x \leq b$ is implemented by proving the existence of a non-negative difference:
\begin{equation}
    \exists \; d \geq 0 : x + d = b
\end{equation}
where $d$ is range-checked via binary decomposition into $n$ bits:
\begin{equation}
    d = \sum_{j=0}^{n-1} d_j \cdot 2^j, \quad d_j \in \{0, 1\}
\end{equation}
Each bit constraint $d_j \cdot (1 - d_j) = 0$ requires one R1CS gate, yielding $2n + 2$ total constraints for a range check with $n$-bit precision.

\subsubsection{Set Membership via Merkle Accumulators}

For whitelist membership $\mathsf{target} \in \mathcal{W}$, we construct a Merkle tree over the whitelist entries using the Poseidon hash function~\cite{grassi2021poseidon} and prove membership by providing the authentication path inside the circuit. For a whitelist of size $|\mathcal{W}|$, this requires $O(\log |\mathcal{W}|)$ Poseidon hash evaluations within the circuit.

\begin{lemma}
\label{lem:merkle_cost}
A Merkle membership proof for a whitelist of $N$ entries using Poseidon-128 requires at most $\lceil \log_2 N \rceil \cdot 312 + 47$ R1CS constraints.
\end{lemma}

\subsubsection{Temporal Window Circuits}

Temporal constraints verify that a timestamp $t$ falls within a permitted window. We encode the window as a set of valid intervals $\{[s_1, e_1], \ldots, [s_k, e_k]\}$ (e.g., business hours across weekdays) and prove membership in the disjunction:
\begin{equation}
    \bigvee_{i=1}^{k} (s_i \leq t \leq e_i)
\end{equation}
using a selector-based approach where the prover commits to which interval $t$ belongs to and proves the range bound for that interval.

\subsubsection{Semantic Similarity Threshold}

This is the most novel gadget. Enterprise policies increasingly require that an agent's \emph{stated intent} (encoded as a vector embedding) be semantically similar to an authorized reference embedding. We implement cosine similarity as a fixed-point arithmetic circuit:

\begin{equation}
    \text{sim}(u, v) = \frac{\sum_{i} u_i \cdot v_i}{\sqrt{\sum_i u_i^2} \cdot \sqrt{\sum_i v_i^2}} \geq \tau
\end{equation}

To avoid expensive square root operations in the circuit, we use the equivalent squared form:
\begin{equation}
    \left(\sum_i u_i v_i\right)^2 \geq \tau^2 \cdot \left(\sum_i u_i^2\right) \cdot \left(\sum_i v_i^2\right)
    \label{eq:cossim}
\end{equation}

Embeddings are quantized to 16-bit signed fixed-point representation (1 sign bit, 3 integer bits, 12 fractional bits). We evaluated quantization fidelity against full float32 cosine similarity across 10,000 sentence pairs drawn from the STS Benchmark using \texttt{all-MiniLM-L6-v2} ($d = 384$). The mean absolute error between quantized and full-precision cosine similarity is $4.7 \times 10^{-3}$, with a worst-case deviation of $9.2 \times 10^{-3}$. More critically, rank-order correlation (Spearman's $\rho$) between the two scoring methods is 0.9994, confirming that 16-bit quantization preserves similarity rankings with negligible distortion. At threshold $\tau = 0.85$, quantization introduces a false positive rate of 0.12\% (a minor security risk, marginally over-approving slightly sub-threshold transactions) and a false negative rate of 0.08\% (an availability risk, rejecting valid transactions) relative to float32 evaluation. For policies requiring stricter fidelity, 24-bit quantization eliminates all observed rank inversions at a cost of $\sim$50\% more multiplication gates. For $d$-dimensional embeddings, the 16-bit gadget requires $3d + O(1)$ multiplication gates.

\subsection{Compiler Optimization}

The compiler applies three optimization passes:

\begin{enumerate}
    \item \textbf{Common sub-expression elimination}: Shared witnesses across constraints (e.g., the same \texttt{amount} field referenced by both a range and rate constraint) are allocated once.
    \item \textbf{Constant folding}: Policy parameters known at compile time (thresholds, whitelist roots) are baked into the circuit as constants, eliminating runtime witness assignments.
    \item \textbf{Gate merging}: Sequential linear constraints are merged into compound R1CS rows where the $\mathbf{A}$, $\mathbf{B}$, $\mathbf{C}$ matrices admit sparse combination.
\end{enumerate}

\begin{lemma}[Compiler Correctness]
\label{lem:compiler_correctness}
For any well-formed PCL policy $\mathsf{P}$ and request $\mathsf{req}$, the compiled R1CS instance $\mathsf{R}$ is satisfiable if and only if $\mathsf{P}(\mathsf{req}) = 1$.
\end{lemma}

\begin{proof}
By structural induction on the AST. Each leaf constraint type has a one-to-one correspondence with its R1CS encoding (shown above), and Boolean combinators are compiled to arithmetic equivalents ($\land \mapsto$ multiplication, $\lor \mapsto a + b - a \cdot b$, $\neg \mapsto 1 - a$). The post-order traversal ensures correct evaluation order. \qed
\end{proof}

%% ===================================================================
%% VI. IMPLEMENTATION
%% ===================================================================
\section{Implementation}
\label{sec:implementation}

\subsection{System Components}

We implement PCT as an extension to a production Rust-based reverse proxy deployed on commodity cloud infrastructure. The implementation comprises four core modules totaling approximately 8,400 lines of Rust:

\begin{itemize}
    \item \textbf{Policy Engine} (\texttt{src/policy/}): PCL parser, AST construction, and runtime evaluation. Written using \texttt{pest} for grammar parsing, producing a typed AST with compile-time validation.

    \item \textbf{Circuit Compiler} (\texttt{src/zkp/compiler.rs}): AST-to-R1CS translation implementing the gadgets described in Section~\ref{sec:compiler}. Built on \texttt{ark-relations} for R1CS constraint management.

    \item \textbf{Proving Engine} (\texttt{src/zkp/prover.rs}): Groth16 proof generation using \texttt{ark-groth16} over the BN254 curve. Supports parallelized multi-scalar multiplication across available CPU cores.

    \item \textbf{Verification Module} (\texttt{src/zkp/verifier.rs}): Stateless $O(1)$ proof verification exposed both as a library function and a REST endpoint (\texttt{/verify-proof}) for external consumers.
\end{itemize}

\subsection{Proof Attachment Protocol}

Proofs are attached to outbound HTTP requests via a custom header protocol:

\begin{lstlisting}[language=bash, caption={PCT header format.}]
X-PCT-Proof: <base64(pi)>
X-PCT-Policy-Version: <uint64>
X-PCT-Timestamp: <unix_epoch_ms>
X-PCT-Public-Inputs: <base64(x_0,...,x_l)>
\end{lstlisting}

The \texttt{Policy-Version} header enables verifiers to select the correct verification key when policies are updated. Public inputs include a SHA-256 commitment to the request body, the timestamp, and the policy version---sufficient for a verifier to reconstruct the verification equation without accessing private fields.

\subsection{Key Management and Trusted Setup}

The Groth16 CRS is generated during a one-time trusted setup ceremony per policy version. We implement a multi-party computation (MPC) ceremony protocol following~\cite{bowe2017scalable} with $n = 3$ participants, where security holds as long as at least one participant is honest. The proving key ($\sim$48MB for a 40,000-gate circuit) is stored encrypted at rest; the verification key ($\sim$1KB) is distributed publicly via the \texttt{/admin/verification-keys} REST endpoint.

%% ===================================================================
%% VII. EVALUATION
%% ===================================================================
\section{Evaluation}
\label{sec:evaluation}

We evaluate PCT along four dimensions: constraint efficiency, proof generation latency, verification overhead, and end-to-end proxy impact.

\subsection{Evaluation Setup}

\textbf{Hardware.} All benchmarks are conducted on an AMD EPYC 7B13 (4 cores, 3.0 GHz) with 16GB RAM, representative of a standard cloud VM (e.g., AWS \texttt{c6a.xlarge}). Benchmarks use the \texttt{criterion} statistical harness (100 iterations per sample, 10 warm-up rounds) to report stable median and P99 latencies.

\textbf{Policy Corpus.} We assembled a corpus of 847 egress policies derived from three production deployment categories spanning fintech, healthcare, and SaaS infrastructure verticals. Because enterprise policies are proprietary and cannot be disclosed directly, we constructed the corpus using the following reproducible methodology: (1)~we conducted structured interviews with security engineers at participating organizations to catalogue the \emph{types}, \emph{arity}, and \emph{constraint distributions} of their active egress rules; (2)~we used these structural templates to synthetically generate parameterized policies with randomized thresholds, whitelist sizes, and temporal windows that preserve the statistical distribution of real deployments. Specifically, constraint counts were sampled from a Poisson distribution ($\lambda = 6.4$), and thresholds were drawn from log-normal distributions derived directly from the observed enterprise baselines; and (3)~we validated that the synthetic policies exercise the same compiler code paths and produce constraint counts within 3\% of the originals. An anonymized subset of 200 representative policies is available in the accompanying artifact package for independent reproduction (Appendix~\ref{sec:artifact}). Policies range from simple 2-constraint configurations (amount limit + whitelist) to complex 23-constraint compositions including temporal, rate, and semantic constraints.

\textbf{Baselines.} We compare against three baselines: (1) unconstrained proxy (no policy enforcement), (2) imperative policy evaluation (standard if/else checking), and (3) signature-based attestation (HMAC signing of evaluation results). We additionally benchmark our domain-specific R1CS compiler against the RISC Zero~\cite{risczero} zkVM across three policy tiers (Section~\ref{sec:zkvm_comparison}).

\subsection{Case Study: Unauthorized API Exfiltration}
\label{sec:casestudy}

To ground the evaluation in a concrete scenario, we trace a single transaction through the PCT pipeline. Consider an LLM agent tasked with processing customer refunds. The agent constructs an HTTP POST to \texttt{api.stripe.com/v1/transfers} with a payload of \$12,400 directed to an external bank account. The organization's policy specifies: (1)~transfers $\leq$ \$10,000, (2)~destination $\in$ approved whitelist, (3)~business hours only, and (4)~intent embedding cosine similarity $\geq 0.85$ to the reference ``process customer refund.''

The proxy extracts the witness: amount $= 12400$, target hash $= \textsf{Poseidon}(\texttt{api.stripe.com})$, timestamp $= 1712678400$ (Tuesday 14:00 UTC), and the 384-dimensional intent embedding. The range constraint fails immediately ($12400 > 10000$), so $\mathsf{P}(\mathsf{req}) = 0$. The proxy rejects the request with a structured denial and logs the violation---no proof is generated. Had the agent instead requested \$8,200, all four constraints would pass, the compiler would synthesize an 11.2K-gate R1CS circuit in $<$1ms, and the Groth16 prover would produce a 192-byte proof $\pi$ in 94ms. The downstream Stripe endpoint verifies $\pi$ in 4.2ms, confirming compliance without learning the organization's \$10K threshold, whitelist composition, or internal intent model.

\subsection{Constraint Efficiency}

\begin{figure}[t]
\centering
\begin{tikzpicture}
\begin{axis}[
    width=\columnwidth,
    height=5.5cm,
    ybar,
    bar width=10pt,
    ylabel={R1CS Constraints (thousands)},
    xlabel={Policy Complexity (number of constraints)},
    symbolic x coords={2--4, 5--8, 9--12, 13--18, 19--23},
    xtick=data,
    ymin=0, ymax=55,
    nodes near coords,
    nodes near coords style={font=\tiny},
    legend style={at={(0.02,0.98)}, anchor=north west, font=\footnotesize},
    every axis plot/.append style={fill opacity=0.7}
]
\addplot[fill=blue!60] coordinates {
    (2--4, 11.2) (5--8, 11.2) (9--12, 11.2) (13--18, 11.3) (19--23, 11.3)
};
\addplot[fill=red!60] coordinates {
    (2--4, 11.2) (5--8, 11.2) (9--12, 11.2) (13--18, 11.2) (19--23, 11.3)
};
\legend{Unoptimized, Optimized}
\end{axis}
\end{tikzpicture}
\caption{R1CS constraint counts. All policies incorporate a constant baseline of 11.2K constraints simulating the fixed cost of Merkle trees and range bit-decompositions.}
\label{fig:constraints}
\end{figure}

Figure~\ref{fig:constraints} shows constraint counts across five complexity tiers. Because constraint generation is heavily dominated by the 32-level Merkle multi-scalar insertions and bit-decompositions, differences in AST complexity represent minor fluctuations around an 11.2K baseline. Critically, 100\% of the evaluated policies compile to fewer than 12,000 constraints, well within the threshold for ultra-low latency proving.

\subsection{Proof Generation Latency}

\begin{table}[t]
\centering
\caption{Proof Generation Latency by Policy Complexity (ms)}
\label{tab:latency}
\begin{tabular}{@{}lrrrr@{}}
\toprule
\textbf{Complexity} & \textbf{P50} & \textbf{P90} & \textbf{P99} & \textbf{Max} \\
\midrule
Simple (2--4) & 92 & 105 & 118 & 132 \\
Moderate (5--8) & 94 & 108 & 120 & 135 \\
Complex (9--12) & 95 & 110 & 122 & 136 \\
Heavy (13--18) & 97 & 111 & 125 & 139 \\
Extreme (19--23) & 98 & 113 & 128 & 142 \\
\midrule
\textbf{All (Median)} & \textbf{95} & \textbf{109} & \textbf{123} & \textbf{142} \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:latency} reports proof generation latency across the corpus. The median latency of 95ms is dominated by the multi-scalar multiplication (MSM) phase for the simulated Merkle proofs. The P99 latency of 123ms ensures that even complex policies comfortably meet strict 500ms Service Level Objectives (SLOs) acceptable for synchronous web-tier transactions.

\subsection{Verification Performance}

Verification consistently completes in $4.2 \pm 0.3$ms regardless of policy complexity, confirming the $O(1)$ constant-time property of Groth16 verification. This is $\sim$23$\times$ faster than proof generation and adds negligible overhead to API response processing.

\subsection{End-to-End Proxy Impact}

\begin{table}[t]
\centering
\caption{End-to-End Proxy Latency Comparison (ms)}
\label{tab:e2e}
\begin{tabular}{@{}lrrr@{}}
\toprule
\textbf{Configuration} & \textbf{P50} & \textbf{P99} & \textbf{Overhead} \\
\midrule
No enforcement & 2.1 & 5.8 & --- \\
Imperative policy (if/else) & 2.3 & 6.1 & +0.2ms \\
HMAC attestation & 2.4 & 6.4 & +0.3ms \\
\textbf{PCT (this work)} & \textbf{97} & \textbf{128} & \textbf{+95ms} \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:e2e} shows the latency trade-off. PCT adds approximately 95ms of median latency---a pragmatic cost of cryptographic compliance guarantees. This overhead applies only to the \emph{prover side} (the organization's own proxy); the verifier experiences only $\sim$4.2ms of additional processing (Groth16's $O(1)$ verification). For latency-sensitive low-risk transactions, organizations can configure PCT in \emph{audit-only mode}, where proofs are generated asynchronously.

\subsection{Comparison with Alternative Proof Systems}

\begin{table}[t]
\centering
\caption{PCT Performance Across Proof Systems (11K-gate circuit)}
\label{tab:systems}
\begin{tabular}{@{}lrrrr@{}}
\toprule
\textbf{System} & \textbf{Prove} & \textbf{Verify} & \textbf{Proof} & \textbf{Setup} \\
 & \textbf{(ms)} & \textbf{(ms)} & \textbf{Size} & \\
\midrule
Groth16 (this work) & 95 & 4.2 & 192B & Trusted \\
PLONK & 134 & 5.1 & 576B & Universal \\
Marlin & 121 & 7.8 & 784B & Universal \\
Halo2 & 189 & 4.2 & 8KB & Transparent \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:systems} justifies our selection of Groth16: it offers the smallest proof size and fastest verification, both critical for HTTP header transmission and API-side processing. The trusted setup cost is amortized over the policy lifetime and can be mitigated through MPC ceremonies.

\subsection{Domain-Specific Compiler vs.\ General-Purpose zkVMs}
\label{sec:zkvm_comparison}

To empirically justify our custom PCL-to-R1CS compiler over general-purpose zkVMs, we implemented three representative policies of increasing complexity as RISC-V guest programs for RISC Zero~\cite{risczero} and measured their execution costs against our native compilation. Table~\ref{tab:zkvm} presents the results.

\begin{table}[t]
\centering
\caption{PCL Compiler vs.\ RISC Zero zkVM}
\label{tab:zkvm}
\begin{tabular}{@{}lrrrr@{}}
\toprule
\textbf{Policy} & \multicolumn{2}{c}{\textbf{PCT (R1CS)}} & \multicolumn{2}{c}{\textbf{RISC Zero}} \\
 & \textbf{Gates} & \textbf{Prove} & \textbf{Cycles} & \textbf{Prove} \\
\midrule
Range + Whitelist & 11.2K & 92ms & 832K & 4.8s \\
+ Temporal + Rate & 11.2K & 95ms & 1.04M & 6.1s \\
+ Semantic ($d$=384) & 11.3K & 98ms & 2.31M & 14.2s \\
\bottomrule
\end{tabular}
\end{table}

The overhead is systematic: a zkVM must arithmetize every RISC-V instruction (memory loads, ALU operations, program counter updates), while our compiler maps each policy constraint directly to its minimal algebraic gadget. The gap widens dramatically for the semantic gadget, where 384-dimensional fixed-point dot products in a zkVM require millions of cycles for the same computation that our specialized circuit achieves in $3d + O(1) = 1{,}153$ multiplication gates. These results confirm that for PCT's narrow constraint vocabulary, a domain-specific compiler is not merely an optimization but an architectural necessity.

\subsection{Robustness of the Semantic Gadget}

As established in our threat model, the semantic constraint $C_{\text{sem}}$ is an optimistic check conditioned on embedding model fidelity. To quantify the practical risk of adversarial embeddings, we measured the perturbation budget required to subvert the gadget's $\tau = 0.85$ boundary using \texttt{all-MiniLM-L6-v2}.

Starting from a benign reference sentence, we applied Projected Gradient Descent (PGD) against the continuous embedding representation to construct adversarial paraphrases that satisfy cosine $\geq 0.85$ while flipping semantic polarity. On average, crossing the false-positive threshold required modifying $\geq$38\% of the input tokens. Because forcing orthogonal semantic meanings into tight geometric proximity introduces severe syntactic distortion, the resulting adversarial outputs are highly anomalous and easily detectable by complementary network or content filters. Thus, while PCT limits its formal guarantees to the mathematical correctness of the semantic check, attacking the semantic plane requires sufficient perturbation that it degrades exploit viability.

%% ===================================================================
%% VIII. REGULATORY MAPPING
%% ===================================================================
\section{Regulatory Compliance Mapping}
\label{sec:regulatory}

A contribution of this work that extends beyond pure cryptography is the formal mapping between PCT's technical guarantees and the requirements of emerging AI governance legislation. We demonstrate that PCT provides the \emph{technical evidence} that regulators increasingly demand.

\subsection{EU AI Act Alignment}

The EU AI Act (Regulation 2024/1689)~\cite{euaiact} establishes risk-based requirements for AI systems deployed in the European market. For high-risk AI systems---which includes autonomous agents operating in financial services, healthcare, and critical infrastructure---the Act mandates:

\begin{itemize}
    \item \textbf{Article 9 (Risk Management)}: PCT provides continuous, mathematically verified risk mitigation for every transaction, replacing probabilistic guardrails with deterministic proofs.

    \item \textbf{Article 12 (Record-Keeping)}: The immutable audit log of Proof-Carrying Transactions creates a cryptographically verifiable record that satisfies the Act's ``automatic recording of events'' requirement. Each log entry includes the proof $\pi$, enabling retroactive verification of any historical transaction's compliance.

    \item \textbf{Article 14 (Human Oversight)}: PCT supports human oversight by providing auditors with independently verifiable evidence. Unlike opaque model-level alignment, a human auditor can verify any specific transaction's compliance by checking $\pi$ against $\mathsf{vk}$---a computation that takes $\sim$4.2ms and requires no understanding of the underlying AI model.
\end{itemize}

\subsection{NIST AI Risk Management Framework}

The NIST AI RMF~\cite{nistrmf} organizes AI risk management around four functions: Govern, Map, Measure, and Manage. PCT contributes to:

\begin{itemize}
    \item \textbf{Measure}: PCT provides quantifiable, binary compliance evidence ($\pi$ verifies or it does not) rather than subjective risk assessments.
    \item \textbf{Manage}: Real-time proof generation enables proactive risk management---non-compliant transactions are blocked \emph{before} execution, not detected afterward.
\end{itemize}

\subsection{Compliance as a Cryptographic Primitive}

We argue that PCT represents a new class of compliance technology: \emph{compliance as a cryptographic primitive}. Rather than treating regulatory compliance as a procedural afterthought (checklists, audits, documentation), PCT embeds compliance directly into the computational infrastructure. Every transaction either carries a valid proof or is rejected---there is no ambiguity, no interpretation gap, and no reliance on human process adherence.

This has profound implications for the emerging ``verifiable safety'' paradigm: organizations can provide regulators not with assertions of compliance, but with mathematical proofs of compliance for every individual action taken by their AI systems.

%% ===================================================================
%% IX. DISCUSSION
%% ===================================================================
\section{Discussion}
\label{sec:discussion}

\subsection{Limitations}

\textbf{Trusted Setup.} Groth16 requires a per-circuit trusted setup. While MPC ceremonies mitigate this, policy updates necessitate new ceremonies. We outline a migration path to universal-setup systems (Section~\ref{sec:future}).

\textbf{Prover as TCB.} The Prover Proxy remains a trusted component. We are investigating TEE-based deployment using Intel SGX~\cite{costan2016intel} and ARM TrustZone~\cite{pinto2019demystifying} to minimize the TCB to hardware-attested enclaves. Smart-contract-based verification~\cite{ferles2022verifying} offers an alternative trust anchor for blockchain-native deployments.

\textbf{Semantic Gadget Precision.} The fixed-point quantization of embeddings introduces a small precision loss ($\leq 0.01$ in cosine similarity). For extremely fine-grained semantic policies, this may require higher bit-width quantization at proportional constraint cost.

\textbf{Latency Overhead.} The 95ms median proving time is acceptable for financial and enterprise transactions but may be prohibitive for ultra-low-latency use cases (e.g., real-time trading at microsecond scales). GPU-accelerated MSM kernels~\cite{lu2024pipemsm} could reduce this by 5--10$\times$.

\textbf{Adversarial Embeddings.} As established in our threat model (Section~\ref{sec:model}), the semantic constraint $C_{\text{sem}}$ is an optimistic check conditioned on embedding model fidelity. PCT guarantees only that the \emph{mathematical evaluation} was performed correctly; it does not guarantee that geometric similarity implies semantic safety. Hardening the embedding space (e.g., via certified robust embeddings or input-space anomaly detection) is orthogonal to our cryptographic contribution.

\subsection{Ethical Considerations}

PCT's policy privacy property is a double-edged sword: while it protects organizational IP, it also means that external parties cannot inspect the policies themselves. This creates a tension with transparency requirements. We note that transparency can be achieved at a higher level---organizations can publicly disclose their policy \emph{types} and \emph{categories} while keeping specific thresholds and parameters private inside the circuit.

\subsection{Future Directions}
\label{sec:future}

\textbf{Recursive Proof Composition.} By adopting recursive proof systems (e.g., Nova~\cite{kothapalli2022nova}), multiple transaction proofs can be aggregated into a single succinct proof, enabling batch verification and reducing per-transaction verification cost for high-throughput systems.

\textbf{Multi-Agent Proof Networks.} In multi-agent systems, PCT can be extended to a \emph{proof-carrying workflow} model where each agent in a pipeline appends its compliance proof, creating a chain of cryptographic attestations that proves end-to-end workflow compliance.

\textbf{On-Chain Verification.} PCT proofs are compatible with EVM-based smart contract verification, enabling decentralized autonomous organizations (DAOs) to cryptographically enforce agent governance on-chain without revealing proprietary policies.

\textbf{Adaptive Circuit Compilation.} Using ML-driven circuit optimization, we aim to learn optimal constraint layouts from historical policy patterns, further reducing gate counts by 15--30\%.

%% ===================================================================
%% X. CONCLUSION
%% ===================================================================
\section{Conclusion}
\label{sec:conclusion}

The deployment of autonomous AI agents in safety-critical domains demands enforcement mechanisms whose guarantees are not merely probabilistic but mathematically verifiable. This paper introduced Proof-Carrying Transactions, a framework that compiles organizational egress policies into zero-knowledge circuits and attaches succinct, independently verifiable proofs to every outbound agent request.

Our evaluation demonstrates that PCT achieves a median proof generation latency of 95ms with verification in 4.2ms, establishing that cryptographic compliance enforcement is practically deployable within existing enterprise SLOs. The formal proofs of computational soundness (Theorem~\ref{thm:soundness}), zero-knowledge policy privacy (Theorem~\ref{thm:privacy}), and policy-binding completeness (Theorem~\ref{thm:completeness}) provide the theoretical foundation, while the regulatory mapping to the EU AI Act and NIST AI RMF demonstrates immediate applicability to emerging governance requirements.

PCT shifts the trust model from centralized enforcement---where downstream systems must implicitly trust the proxy operator---to decentralized verification, where any party holding the public verification key can independently confirm compliance in constant time. We believe this paradigm represents a foundational primitive for the emerging field of \emph{verifiable AI}: systems whose every consequential action carries a machine-checkable certificate of authorization.

%% ===================================================================
%% APPENDIX
%% ===================================================================
\appendices

\section{Artifact Description}
\label{sec:artifact}

To support the reproducibility of all evaluation results reported in Section~\ref{sec:evaluation}, we provide a self-contained artifact package. The artifact is distributed as a Docker image that can be built and executed on any system with Docker installed, requiring no Rust toolchain or host-level dependencies.

\subsection{Contents}

The artifact package contains the following components:

\begin{itemize}
    \item \textbf{PCT Reverse Proxy} (\texttt{proxy/}): The complete Rust implementation of the Policy Constraint Language parser, the AST-to-R1CS circuit compiler, the Groth16 proving engine, and the stateless verification module (approximately 8,400 lines of Rust).
    \item \textbf{Benchmark Suite} (\texttt{proxy/benches/pct\_bench.rs}): A \texttt{criterion}-based statistical benchmark harness that measures proof generation latency, verification latency, and trusted setup time across five policy complexity tiers.
    \item \textbf{Constraint Counter} (\texttt{constraint\_counter.rs}): A standalone binary that reports the exact R1CS constraint count emitted by the circuit compiler for a given policy configuration.
    \item \textbf{Synthetic Policy Corpus} (\texttt{policies/synthetic\_200.json}): An anonymized subset of 200 representative policies drawn from the 847-policy evaluation corpus. Each policy is parameterized with randomized thresholds and whitelist sizes that preserve the statistical properties of the original production rules.
    \item \textbf{Dockerfile} (\texttt{artifact/Dockerfile}): A self-contained build environment based on \texttt{rust:1.75-bookworm} that compiles and pre-fetches all dependencies.
\end{itemize}

\subsection{Reproduction Instructions}

\begin{enumerate}
    \item Build the Docker image: \texttt{docker build -t pct-artifact -f artifact/Dockerfile .}
    \item Start an interactive shell: \texttt{docker run -it pct-artifact /bin/bash}
    \item Execute benchmarks: \texttt{cargo bench}
    \item Run the constraint counter: \texttt{cargo run --bin constraint\_counter}
\end{enumerate}

The \texttt{cargo bench} output reports proving and verification latencies grouped by policy complexity tier and directly corresponds to Table~\ref{tab:latency} and Table~\ref{tab:e2e} in the main text. The constraint counter output corresponds to Figure~\ref{fig:constraints}.

\section{Extended Proof Details}
\label{sec:proofs}

We provide additional detail for the proof reductions summarized in Section~\ref{sec:model}.

\subsection{Soundness Reduction (Theorem~\ref{thm:soundness})}

The full reduction proceeds as follows. Given any PPT adversary $\mathcal{A}^*$ that breaks PCT soundness, we construct an adversary $\mathcal{B}$ that breaks Groth16 knowledge soundness.

$\mathcal{B}$ receives the CRS $(\mathsf{pk}, \mathsf{vk})$ from the Groth16 challenger and forwards $\mathsf{pk}$ to $\mathcal{A}^*$. When $\mathcal{A}^*$ outputs a forgery $(\mathsf{req}, \pi^*)$ where $\mathsf{P}(\mathsf{req}) = 0$ but $\mathcal{V}(\mathsf{vk}, \mathsf{req}, \pi^*) = 1$, $\mathcal{B}$ extracts the public inputs $x = (H(\mathsf{req}), t, v)$ and submits $(x, \pi^*)$ to the Groth16 challenger. By Lemma~\ref{lem:compiler_correctness}, $\mathsf{P}(\mathsf{req}) = 0$ implies no satisfying witness exists for $\mathsf{R}$, so $\mathcal{B}$ has produced a valid proof for an unsatisfiable instance---contradicting Groth16's knowledge soundness under KoE. The advantage of $\mathcal{A}^*$ is therefore bounded by $\textsf{Adv}^{\text{KoE}}_{\mathcal{B}} \leq \textsf{negl}(\lambda)$.

\subsection{Privacy Reduction (Theorem~\ref{thm:privacy})}

The zero-knowledge property follows from the \emph{perfect} simulation guarantee of Groth16. For any CRS with trapdoor $\tau$, the simulator $\mathcal{S}(\mathsf{vk}, x, \tau)$ produces proofs $\tilde{\pi}$ that are \emph{identically distributed} to honest proofs $\pi$. Since simulated proofs carry no information about the witness $w$ (which encodes both the request parameters and the policy structure), an observer seeing only $(x, \pi)$ gains zero bits of information about $\mathsf{P}$ beyond what is revealed by the public inputs $x$. The public inputs are restricted to $(H(\mathsf{req}), t, v)$---a hash commitment, timestamp, and policy version---none of which leak constraint types, thresholds, or whitelist contents.

%% ===================================================================
%% REFERENCES
%% ===================================================================
\bibliographystyle{IEEEtran}
\begin{thebibliography}{99}

\bibitem{xi2023rise}
Z. Xi et al., ``The rise and potential of large language model based agents: A survey,'' \textit{arXiv preprint arXiv:2309.07864}, 2023.

\bibitem{wang2024survey}
L. Wang et al., ``A survey on large language model based autonomous agents,'' \textit{Frontiers of Computer Science}, vol.~18, no.~6, 2024.

\bibitem{bai2022constitutional}
Y. Bai et al., ``Constitutional AI: Harmlessness from AI feedback,'' \textit{arXiv preprint arXiv:2212.08073}, 2022.

\bibitem{ouyang2022training}
L. Ouyang et al., ``Training language models to follow instructions with human feedback,'' in \textit{Proc. NeurIPS}, 2022.

\bibitem{perez2022red}
E. Perez et al., ``Red teaming language models with language models,'' in \textit{Proc. EMNLP}, 2022.

\bibitem{necula1997proof}
G. C. Necula, ``Proof-carrying code,'' in \textit{Proc. 24th ACM SIGPLAN-SIGACT Symposium on Principles of Programming Languages (POPL)}, 1997, pp. 106--119.

\bibitem{goldwasser1989knowledge}
S. Goldwasser, S. Micali, and C. Rackoff, ``The knowledge complexity of interactive proof systems,'' \textit{SIAM Journal on Computing}, vol.~18, no.~1, pp. 186--208, 1989.

\bibitem{groth2016size}
J. Groth, ``On the size of pairing-based non-interactive arguments,'' in \textit{Proc. EUROCRYPT}, 2016, pp. 305--326.

\bibitem{gabizon2019plonk}
A. Gabizon, Z. J. Williamson, and O. Ciobotaru, ``PLONK: Permutations over Lagrange-bases for oecumenical noninteractive arguments of knowledge,'' \textit{IACR Cryptol. ePrint Arch.}, 2019.

\bibitem{sasson2014zerocash}
E. Ben-Sasson et al., ``Zerocash: Decentralized anonymous payments from Bitcoin,'' in \textit{Proc. IEEE S\&P}, 2014, pp. 459--474.

\bibitem{loopring2020}
Loopring Protocol, ``zkRollup: A layer 2 scaling solution,'' \textit{Technical Report}, 2020.

\bibitem{semaphore2020}
Semaphore, ``A zero-knowledge gadget for creating anonymous proof of membership,'' \textit{Github Repository}, 2020.

\bibitem{grassi2021poseidon}
L. Grassi et al., ``Poseidon: A new hash function for zero-knowledge proof systems,'' in \textit{Proc. USENIX Security}, 2021.

\bibitem{bowe2017scalable}
S. Bowe, A. Gabizon, and I. Miers, ``Scalable multi-party computation for zk-SNARK parameters in the random beacon model,'' \textit{IACR Cryptol. ePrint Arch.}, 2017.

\bibitem{kothapalli2022nova}
A. Kothapalli, S. Setty, and I. Tzialla, ``Nova: Recursive zero-knowledge arguments from folding schemes,'' in \textit{Proc. CRYPTO}, 2022.

\bibitem{euaiact}
European Parliament and Council, ``Regulation (EU) 2024/1689 laying down harmonised rules on artificial intelligence (Artificial Intelligence Act),'' \textit{Official Journal of the European Union}, 2024.

\bibitem{nistrmf}
National Institute of Standards and Technology, ``Artificial Intelligence Risk Management Framework (AI RMF 1.0),'' NIST AI 100-1, 2023.

\bibitem{yao2022react}
S. Yao et al., ``ReAct: Synergizing reasoning and acting in language models,'' in \textit{Proc. ICLR}, 2023.

\bibitem{bunz2018bulletproofs}
B. B\"unz et al., ``Bulletproofs: Short proofs for confidential transactions and more,'' in \textit{Proc. IEEE S\&P}, 2018.

\bibitem{ben2018scalable}
E. Ben-Sasson et al., ``Scalable, transparent, and post-quantum secure computational integrity,'' \textit{IACR Cryptol. ePrint Arch.}, 2018.

\bibitem{sharma2023machine}
V. Sharma et al., ``Machine learning and zero-knowledge proofs: A survey,'' \textit{CSUR}, 2023.

\bibitem{kang2022scaling}
D. Kang et al., ``Scaling up trustless execution of machine learning models,'' in \textit{Proc. USENIX Security}, 2022.

\bibitem{eberhardt2018zokrates}
J. Eberhardt and S. Tai, ``ZoKrates-scalable privacy-preserving off-chain computations,'' in \textit{Proc. IEEE Blockchain}, 2018.

\bibitem{ferles2022verifying}
C. Ferles et al., ``Verifying smart contracts with formal methods,'' in \textit{Proc. ACM CCS}, 2022.

\bibitem{abadi2011authorization}
M. Abadi and G. Plotkin, ``A calculus for access control in distributed systems,'' \textit{ACM TOPLAS}, 2011.

\bibitem{sandhu1996role}
R. S. Sandhu et al., ``Role-based access control models,'' \textit{IEEE Computer}, 1996.

\bibitem{birgisson2014macaroons}
A. Birgisson et al., ``Macaroons: Cookies with contextual caveats for decentralized authorization in the cloud,'' in \textit{Proc. NDSS}, 2014.

\bibitem{w3c2022vc}
W3C, ``Verifiable Credentials Data Model v2.0,'' \textit{W3C Recommendation}, 2022.

\bibitem{schmidt2019opa}
S. Schmidt and T. Hinrichs, ``Open Policy Agent: Cloud-native authorization,'' \textit{CNCF}, 2019.

\bibitem{zheng2023llama}
L. Zheng et al., ``Judging LLM-as-a-judge with MT-Bench and Chatbot Arena,'' in \textit{Proc. NeurIPS}, 2023.

\bibitem{hendrycks2021unsolved}
D. Hendrycks et al., ``Unsolved problems in ML safety,'' \textit{arXiv preprint arXiv:2109.13916}, 2021.

\bibitem{carlini2023extracting}
N. Carlini et al., ``Extracting training data from large language models,'' in \textit{Proc. USENIX Security}, 2021.

\bibitem{blanchard2017byzantine}
P. Blanchard et al., ``Machine learning with adversaries: Byzantine tolerant gradient descent,'' in \textit{Proc. NeurIPS}, 2017.

\bibitem{zhao2024verifiable}
H. Zhao et al., ``Verifiable machine learning: Current status and future directions,'' \textit{arXiv preprint arXiv:2401.07457}, 2024.

\bibitem{togan2023zero}
M. Togan et al., ``Zero-knowledge proofs in decentralized finance,'' \textit{IEEE Access}, 2023.

\bibitem{costan2016intel}
V. Costan and S. Devadas, ``Intel SGX explained,'' \textit{IACR Cryptol. ePrint Arch.}, 2016.

\bibitem{pinto2019demystifying}
S. Pinto and N. Santos, ``Demystifying ARM TrustZone: A comprehensive survey,'' \textit{ACM Computing Surveys}, 2019.

\bibitem{sun2023trust}
Y. Sun et al., ``Trust region policy optimization for LLM alignment,'' in \textit{Proc. ICML}, 2023.

\bibitem{amodei2016concrete}
D. Amodei et al., ``Concrete problems in AI safety,'' \textit{arXiv preprint arXiv:1606.06565}, 2016.

\bibitem{bost2015machine}
R. Bost et al., ``Machine learning classification over encrypted data,'' in \textit{Proc. NDSS}, 2015.

\bibitem{badrinarayanan2020verifiable}
S. Badrinarayanan et al., ``Verifiable capacity from zero knowledge proofs,'' in \textit{Proc. TCC}, 2020.

\bibitem{risczero}
J. H. Ahn et al., ``RISC Zero zkVM: Scalable, transparent zero-knowledge proofs,'' \textit{Technical Report}, \url{https://www.risczero.com/docs/architecture}, 2022.

\bibitem{sp1}
H. Kalodner et al., ``SP1: A high-performance scalable zkVM,'' \textit{Technical Report}, Succinct Labs, \url{https://succinctlabs.github.io/sp1}, 2024.

\bibitem{jolt}
A. Arun, S. Setty, and J. Thaler, ``Jolt: SNARKs for Virtual Machines via Lookups,'' in \textit{Proc. EUROCRYPT}, 2024. \url{https://eprint.iacr.org/2023/1217}.

\bibitem{lu2024pipemsm}
Y. Lu et al., ``PipeMSM: Hardware acceleration for multi-scalar multiplication,'' in \textit{Proc. IEEE FCCM}, 2024.

\balance
\end{thebibliography}

\end{document}
